---
title: "p8105_hw2_TS3175"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Problem 1 
### Read in Mr. Trash Wheel data, omit rows without dumpsterspecific data, and round sports balls to the nearest integer.
```{r}
HH_data = 
  readxl::read_excel("./HW2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names()%>% 
  select(-x15, -x16, -x17)%>% 
  drop_na(dumpster)%>% 
  mutate(sports_balls= as.integer(round(sports_balls, digits=0)))
```
### Read in 2017 precipitation data, omit rows that do not have precipitation data, and create a year variable
```{r}
percip7_data = 
  readxl::read_excel("./HW2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "2017 Precipitation",
  range="A2:B14") %>% 
  janitor::clean_names()%>% 
  drop_na(total)%>% 
  mutate(year=2017)
```
### Read in 2018 precipitation data, omit rows that do not have precipitation data, and create a year variable
```{r}
percip8_data = 
  readxl::read_excel("./HW2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "2018 Precipitation",
  range="A2:B14") %>% 
  janitor::clean_names()%>% 
  drop_na(total)%>% 
  mutate(year=2018)
```
### Combined 2017 and 2018 precipition data 
```{r}
percip_total=
  bind_rows(percip7_data, percip8_data) %>% 
  janitor::clean_names() %>% 
  mutate(month=month.name[month])
```
Data Summary:
The total number of observations in the healthy harbor dataset is `r nrow(HH_data)`. The total number of observations in 2017-2018 precipitation data is `r nrow(percip_total)`. The total precipitation in 2018 was `r sum(pull(percip8_data, total))`. The median sports balls in a dumpster from 2017 is `r median(pull(filter(HH_data, year==2017), sports_balls))`. The key variables in the healthy harbor dataset and the precipitation dataset were total, month, and year.
## Problem 2
### Read in pols month data and create one president variable that encompasses dem and gop. Removed day variable at the end.  
```{r}
pols_month = read_csv("./HW2_data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, into= c("year", "month", "day"),
  convert=TRUE) %>% 
  mutate(month=month.name[month]) %>% 
  mutate(prez_dem=recode(prez_dem, "1"="dem")) %>% 
  mutate(prez_gop=recode(prez_gop, "1" = "gop")) %>%
  pivot_longer(c("prez_dem", "prez_gop"),
               values_to = "president",
               values_drop_na = TRUE) %>% 
   select(-day)
```
## Read in Five Thirty Eight data and arrange data to month day year and remove day variable.
```{r}
snp_data = 
  read_csv("./HW2_data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, into= c("month", "day", "year"),
  convert=TRUE) %>% 
  select(year, month, -day, close)%>% 
  mutate(month=month.name[month])
```
## Read in Unemployment data and tidy data to match previous two data sets.
```{r}
unemp_data=
  read_csv("./HW2_data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(Jan:Dec,
              names_to="month",
              values_to="rate")%>%
  mutate(month=match(month, month.abb)) %>% 
  mutate(month=month.name[month]) %>% 
  mutate(Year=as.integer(Year)) %>% 
  janitor::clean_names()
```
## Merge all datasets
```{r}
Prob_2= pols_month %>% 
  left_join(snp_data,
            by=c("year", "month")) %>% 
  left_join(unemp_data,
            by=c("year", "month"))
```
Data Summary:
The data in the pols-month data set included the the number of national politicians who are democratic or republican for any given month from 1947-2015. With this data, a variable was created categorizing the variable president to democratic and republican polling categories. This dataset has 817 observatins and 10 variables.
The data in the snp dataset is the Standard & Poor's stock market index. The dataset contains the closing values of the stock index of the associated date and the date of the observation for 1950-2015. This dataset contains 787 observations and 3 variables.
The unemployment dataset includes the percentage of unemployment for month of the associated year for 1948-2015. This dataset contains 816 observations and 4 variables.
The three datasets were merged resulting in dataset "prob_2" wich has observations from 1947-2015 and 817 observations and 13 variables. 
The key variables used throughout are month and year.
