---
title: "p8105_hw2_TS3175"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Problem 1 
### Read in Mr. Trash Wheel data, omit rows without dumpsterspecific data, and round sports balls to the nearest integer.
```{r}
HH_data = 
  readxl::read_excel("./HW2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names()%>% 
  select(-x15, -x16, -x17)%>% 
  drop_na(dumpster)%>% 
  mutate(sports_balls= as.integer(round(sports_balls, digits=0)))
```
### Read in 2017 precipitation data, omit rows that do not have precipitation data, and create a year variable
```{r}
percip7_data = 
  readxl::read_excel("./HW2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "2017 Precipitation",
  range="A2:B14") %>% 
  janitor::clean_names()%>% 
  drop_na(total)%>% 
  mutate(year=2017)
```
### Read in 2018 precipitation data, omit rows that do not have precipitation data, and create a year variable
```{r}
percip8_data = 
  readxl::read_excel("./HW2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "2018 Precipitation",
  range="A2:B14") %>% 
  janitor::clean_names()%>% 
  drop_na(total)%>% 
  mutate(year=2018)
```
### Combined 2017 and 2018 precipition data 
```{r}
##Combine precipitation data
percip_total=
  bind_rows(percip7_data, percip8_data) %>% 
  janitor::clean_names() %>% 
  mutate(month=month.name[month])
```
The total number of observations in the healthy harbor dataset is `r nrow(HH_data)`. The total number of observations in 2017-2018 precipitation data is `r nrow(percip_total)`. The total precipitation in 2018 was `r sum(pull(percip_total, total))`. The median sports balls in a dumpster from 2017 is `r median(pull(HH_data, sports_balls))`. The key variables in the healthy harbor dataset and the precipitation dataset were total, month, and year.


